---
title: "9-27-2018_Class - Visualization and EDA 1"
author: "J Shearston"
date: "September 27, 2018"
output: 
  html_document:
    toc: true
---

This R Markdown is a tutorial on ggplots, including different types of plots and some neat options.


```{r setup, include=FALSE}

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

library(tidyverse)
library(ggridges)
```

## Pulling in the NOAA data

```{r pulling noaa data, cache=TRUE}
library(rnoaa)

weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## Start with those plots!


### Scatterplots

####Step 1: Define data (x and y)

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))
```

####Step 2: Add a geom to define what I want to show

Two ways to code a basic scatterplot, one using a pipe.

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()

#or

weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

```

You can save the plot as an object so that you can modify it and play around with it later.

```{r}
plot_weather = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) 

plot_weather + geom_point()
```

*Check -- "plot_weather" is now in the Environment*


####Step 3: Adding fun details, like color and lines

Using the color aesthetic (an aesthetic is part of your data, so in this chunk of code "name" is a variable)

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
```

Adding a smooth curve (geom_smooth) and making the points semi-transparent (alpha). 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```

*Notes: 1) "alpha" is not currently an aesthetic; it doesn't include data but rather alters the appearance of the already present data. 2) "se" is standard error bands*

A bit more on aesthetics...
*The location of the color aesthetic in the code chunk matters. Becuase the color aesthetic has now been moved to the first line, it applies to the entire plot rather than to only the geom_point. Thus, we now have three geometric lines, because the geom_smooth function is applied to the entire plot, and not to a geom_point* 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE)
```

Adding a facet_grid to split out plots by a variable

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

Switching up the data & adding a size aesthetic

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

### Learning Assessment 1

Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in  geom_smooth()).

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>%
  mutate(
    tmax_F = tmax * 1.8 + 32,
    tmin_F = tmin * 1.8 + 32) %>% 
  ggplot(aes(x = tmin_F, y = tmax_F)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

```


### Other Plots

#### Scatter-ish plots

Plot with a fitted line

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) 
```

geom_hex plot and other ways to avoid overplotting

```{r}
library(hexbin)

ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()

ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_bin2d()

ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_density2d()

```


#### Univariate plots

Histograms

```{r}
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()
```

Adjusting histograms - bin width, fill, and "dodge"

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```

Density plot

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
```

*Notes: `Adjust` is similar to `binwidth`, `color` is the outline of each shape.*

Box plot

```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_boxplot()
```

Violin plot

```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  stat_summary(fun.y = median, geom = "point", color = "blue", size = 4)
```

Ridge plot

```{r}
ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```


### Learning Assessment 3 (skipped 2)

*Question*

Make plots that compare precipitation across locations. Try a histogram, a density plot, a boxplot, a violin plot, and a ridgeplot; use aesthetic mappings to make your figure readable.

*Solution*

First a density plot:

```{r}
ggplot(weather_df, aes(x = prcp)) + 
  geom_density(aes(fill = name), alpha = .5) 
```

Next a ridge plot:

```{r}
ggplot(weather_df, aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)
```

Last a boxplot:

```{r}
ggplot(weather_df, aes(y = prcp, x = name)) + 
  geom_boxplot() 
  
```

This is a tough variable to plot because of the highly skewed distribution in each location. Of these, I’d probably choose the boxplot because it shows the outliers most clearly. If the “bulk” of the data were interesting, I’d probably compliment this with a plot showing data for all precipitation less than 100, or for a data omitting days with no precipitation.

```{r}
weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)
```


## Saving and Embedding Plots

Using `ggsave`

```{r}
weather_plot = ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("weather_plot.pdf", weather_plot, width = 8, height = 5)
```

Things to know:

* Figure size in R is controlled by `fig.width`, `fig.height`, and  `fig.asp`. Jeff prefer's a common width and plots that are a little wider than they are tall, so he sets options to fig.width = 6 and  fig.asp = .6

* Figure size inserted into output document is controlled by `out.width` or `out.height`. Jeff likes to have a little padding around the sides of his figures, so he sets out.width = "90%"

* You can set all this using `knitr` at the beginning of the markdown document...but also included just below.

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

